---
title: "Fiji Final Presentation"
author: "Chloe Singer"
date: "2024-12-20"
output: 
  powerpoint_presentation: default
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(leaflet)
library(WDI)
library(forecast)  # Ensure forecast package is loaded for auto.arima
library(dplyr)
library(knitr)  # For tables in Markdown

fiji_gdp <- WDI(country = "FJ", indicator = "NY.GDP.MKTP.CD", start = 2000, end = 2023)
gdp_ts <- ts(fiji_gdp$NY.GDP.MKTP.CD, start = 2000, frequency = 1)
fiji_population <- WDI(country = "FJ", indicator = "SP.POP.TOTL", start = 2000, end = 2023)
fiji_visitors <- WDI(country = "FJ", indicator = "ST.INT.ARVL", start = 2000, end = 2023)
visitor_ts <- ts(fiji_visitors$ST.INT.ARVL, start = 2000, frequency = 1)
```

# Slide 1: Introduction
Fiji, an archipelago of over 330 islands in the South Pacific, is a cultural and economic hub of the region. Known for its stunning natural beauty, Fiji also plays a pivotal role in regional trade and tourism. This presentation delves into Fiji's economic indicators, population trends, visitor statistics, and competitive position among Pacific Island states.

---

# Slide 2: Key Facts About Fiji
- **Location**: Fiji is situated in the South Pacific, northeast of New Zealand.
- **Population**: Approximately 900,000, characterized by a diverse mix of indigenous Fijians, Indo-Fijians, and other ethnic groups.
- **Government**: Parliamentary democracy with a strong emphasis on inclusivity.
- **Economy**:
  - Tourism accounts for over 40% of GDP.
  - Agriculture (primarily sugarcane) and remittances play key roles.
  - Emerging focus on renewable energy and sustainable tourism.

```{r map, echo=FALSE}
leaflet() %>%
  addTiles() %>%
  setView(lng = 178.0650, lat = -17.7134, zoom = 6) %>%
  addMarkers(lng = 178.0650, lat = -17.7134, popup = "Fiji")
```

---

# Slide 3: Economic Trends

## GDP Trends
Fiji's GDP has shown steady growth over the years, with significant contributions from tourism and exports. Efforts to diversify the economy are ongoing.

```{r gdp-trends, echo=FALSE}
plot(gdp_ts, main = "Fiji GDP Over Time", ylab = "GDP (Current US$)", xlab = "Year")
```

## Population Trends
Fiji's population growth reflects increasing urbanization and improvements in healthcare. The government aims to balance urban expansion with environmental sustainability.

```{r population-trends, echo=FALSE}
pop_ts <- ts(fiji_population$SP.POP.TOTL, start = 2000, frequency = 1)
plot(pop_ts, main = "Fiji Population Growth", ylab = "Population", xlab = "Year")
```

---

# Slide 4: Visitor Statistics
Tourism remains the backbone of Fiji's economy. Visitor numbers peaked pre-pandemic and are recovering rapidly as global travel resumes.

- **Key Markets**: Australia, New Zealand, and the United States.
- **Trends**: Growing interest in eco-tourism and cultural tourism.

```{r visitor-trends, echo=FALSE}
plot(visitor_ts, main = "Visitor Trends for Fiji", ylab = "Number of Visitors", xlab = "Year")
```

---

# Slide 5: Comparison with Other Island States

Fiji outperforms many Pacific Island states in GDP, infrastructure, and tourism development. However, challenges such as climate vulnerability and reliance on external markets persist.

```{r comparison, echo=FALSE}
pacific_gdp <- WDI(country = c("FJ", "TO", "VU", "WS"), indicator = "NY.GDP.MKTP.CD", start = 2020, end = 2023) %>%
  rename(GDP = NY.GDP.MKTP.CD) %>%
  group_by(country) %>%
  summarize(GDP = mean(GDP, na.rm = TRUE))

ggplot(pacific_gdp, aes(x = country, y = GDP, fill = country)) +
  geom_bar(stat = "identity") +
  labs(title = "GDP Comparison of Pacific Island States", y = "GDP (Current US$)", x = "Country")
```

---

# Slide 6: SWOT Analysis

- **Strengths**:
  - World-class tourism infrastructure.
  - Abundant natural resources.
  - Strong cultural identity and heritage.
- **Weaknesses**:
  - Overdependence on tourism.
  - Limited diversification in exports.
- **Opportunities**:
  - Expansion into renewable energy.
  - Growth in eco-tourism and high-end hospitality sectors.
- **Threats**:
  - Climate change impacts, particularly rising sea levels.
  - Global economic fluctuations affecting key industries.

---

# Slide 7: Conclusion
Fiji exemplifies resilience and adaptability in a rapidly changing global environment. With strategic investments in sustainable industries and infrastructure, Fiji can secure long-term growth while preserving its natural beauty and cultural legacy.

---

# References
- **World Bank Development Indicators (WDI)**: Comprehensive economic and population data.
- **Fiji Bureau of Statistics**: Visitor trends and demographic insights.
- **Tourism Fiji Reports**: Strategic direction and performance of the tourism sector.

```
